Classless Java

### 2016-03-22 OOPSLA16/sections/FutureWork.tex
little fixes

%\subsection{Qualifiers in Methods} %{Private state
\paragraph{Encapsulation} %{Private state
Since we are using Java8 to encode IB, we are unable to directly support qualifies
like private or protected. In Java 9, where private methods will be allowed in
interfaces; abstract state operations are, however, abstract methods,
so they can not be private.
Still, because the state can only be accessed by methods, it 
is impossible for the user to know if a certain method maps directly to a field
or if it has a default implementation.

@@

Using packages to as a boundary we can make all object interfaces ``with privateness''
invisible types for the rest of the program.
This preserve encapsulation, but limits inheritance:
package protected interfaces can not be seen,
thus can not be implemented from outside the package.

Using Java9 private interface methods
\footnote{Or some ugly pattern available today like public static classes with private methods}
one can define (and inherit) private behaviour; but private state inheritance in IB poses a logical challenge:
How can the heir produce initialization code for a field whose very existence is secret?

@@

In this way

\begin{lstlisting}
@Obj interface MyData{
  private String f();//assuming Java9
  private MyData f(String val);
  }
\end{lstlisting}
Would be expanded into 
\begin{lstlisting}
@Obj interface MyData{
  private Box<String> _f(){return cache(Box<String>.of("hello"));}
  private String f(){return _f().val();}
  private MyData f(String val){return _f().val(val);}
  }
\end{lstlisting}

>>>>

%\subsection{Qualifiers in Methods} %{Private state
\paragraph{Encapsulation} %{Private state
Since we are using Java8 to encode IB, we are unable to directly support qualifies
like private or protected. In Java 9 private methods will be allowed in
interfaces. Abstract state operations are, however, abstract methods,
so they can not be private.
Still, because the state can only be accessed by methods, it 
is impossible for the user to know if a certain method maps directly to a field
or if it has a default implementation.

@@

Using packages as a boundary we can make all object interfaces ``with privateness''
invisible types for the rest of the program.
This preserve encapsulation, but limits inheritance:
package protected interfaces can not be seen,
thus can not be implemented from outside the package.

Using Java9 private interface methods%
\footnote{Or some ugly pattern available today like public static classes with private methods}
one can define (and inherit) private behaviour; but private state inheritance in IB poses a logical challenge:
How can the heir produce initialization code for a field whose very existence is secret?

@@

In this way
\marco{I fixed the code down, was wrong, there was no mention of hello in the original source}
\begin{lstlisting}
@Obj interface Text{
  private String msg();//assuming Java9
  private Text msg(String val);{f("hello");}
  }
\end{lstlisting}
Would be expanded into 
\begin{lstlisting}
@Obj interface Text{
  private Box<String> _msg(){
    return cache(Box<String>.of("hello"));}
  private String msg(){return _msg().val();}
  private Text msg(String val){
    return _msg().val(val);}
  }
\end{lstlisting}

### 2016-03-22 OOPSLA16/sections/Translation.tex
little fixes

To formally characterize the behavior of our annotation and the two levels of guarantees that we offer, we provide some notations and two theorems:
\begin{itemize}
\item We denote with $\C^{\II}$ the name of an interface.
\item An interface table
IT is OK if under such interface table, all interfaces are OK.
\item Since interface tables are just represented as sequences of interfaces we write IT = $\II$ IT' to select a specific interface in a table.
\item IT contains an heir of $\C$ if there is an interface that extends it, or a \Q@new@ that instantiate it.
\end{itemize}

@@

Informally, the theorems mean that for a client program that
typechecks before the translation is applied, if the annotated type has
no subtypes and no objects of that type are created, then type safety
\bruno{of the generated code?}
is guaranteed after the successful translation.

@@

\paragraph{Type preservation:} 
Note that we preferred to introduce self, client and heir coherence instead of referring to conventional
type preservation theorems~\cite{}. The reason is 
to better model how our approach behaves in a object oriented software ecosystem with inheritance,
where only some units may be translated/expanded.
Note that inheritance crucial influence in heir coherence, and how client coherence allows
for only some units to be translated/expanded.\bruno{I don't
  understand the point of the last sentence? Rephrase?}
This means that even a library designer can decide to start using IB
programming and still
retain backward compatibility with its clients.

>>>>

To formally characterize the behavior of our annotation and the two levels of guarantees that we offer, we provide some notations and two theorems:
\begin{itemize}
\item We denote with $\C^{\II}$ the name of an interface.
\item An interface table
IT is OK if under such interface table, all interfaces are OK,
\marcoT{that is, well typed}.
\item Since interface tables are just represented as sequences of interfaces we write IT = $\II$ IT' to select a specific interface in a table.
\item IT contains an heir of $\C$ if there is an interface that extends it, or a \Q@new@ that instantiate it.
\end{itemize}

@@

Informally, the theorems mean that for a client program that
typechecks before the translation is applied, if the annotated type has
no subtypes and no objects of that type are created, then type safety
\marcoT{of the generated code}
is guaranteed after the successful translation.

@@

\paragraph{Type preservation:} 
Note that we preferred to introduce self, client and heir coherence instead of referring to conventional
type preservation theorems~\cite{}. The reason is 
to better model how our approach behaves in a object oriented software ecosystem with inheritance,
where only some units may be translated/expanded.
Note inheritance crucial influence in heir coherence, and how client coherence allows
for only some units to be translated/expanded.
\bruno{I don't
  understand the point of the last sentence? Rephrase?}
  \marco{ok I tried to rephase the next sentence, how is now?}
This means that developers and designers of Java libraries and frameworks
can start using IB (and our \mixin annotation) in the evolution of their products
and still retain backward compatibility with their clients.
